---
title: "SummRt"
output: html_document
date: "2024-09-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# start with individual packages, eventually can just load in SummRt
library(EpiNow2)
library(EpiEstim)
library(rtestim)
library(EpiLPS)
#library(SummRt)
library (tidyverse)
```

Load the dataset we will be fitting the R(t) estimation packages
to in this vignette.

This will eventually use package data that will be documented and
will describe the specific epidemiological use case this
data scenario is meant to represent. 
```{r load-package-data}
# We will eventually replace this with package data specific to the dataset. 
# E.g. this might be our baseline infections, onset, report data
url <- "https://raw.githubusercontent.com/cmilando/RtEval/main/all_data.RDS"
all_data <- readRDS(url(url))
```

Fit each of the packages to the dataset. This will call `SummRt` package
functions which will use S3 methods to wrap around the individual 
R packages. All outputs will be in a standardized format that we can use to
evaluate the R(t) estimation (either from directly comparing to the 
true simulated R(t) or via comparing to an evaluation dataset of held out
retrospective observations).
```{r}

# BAD practice but will source a file
source("02_EPINOW2_infections.R")
```



