---
title: "Epi LPS Rt Vignette"
output: html_document
date: "2024-09-25"
---

# Introduction

This vignette provides an interactive example of **ğ‘…ğ‘¡** estimation from incidence data using the **EpiLPS** package. **EpiLPS** utilizes *Laplacian-P-Splines* to provide rapid, bayesian estimates of the epidemic curve and instantaneous reproduction number among other functionalities.

In this case, we are focusing specifically on the estimation of **ğ‘…ğ‘¡** and demonstrating the use of **summrt** to standardize outputs from **EpiLPS** and related models to a predictable, cleaned format for cross-validation and performance evaluation.

# Prepare the environment

This vignette requires the following libraries:

-   **EpiLPS:** estimates **ğ‘…ğ‘¡**
-   **tidyverse:** clean and format model data to tibbles for plotting
-   **summrt:** standardizes outcome variable formats across epidemiological models

## Install dependencies

```{r}
#install.packages('EpiLPS')
#install.packages("tidyverse")
#install.packages('summrt')
```

## Import libraries

```{r}
require(EpiLPS)
require(tidyverse)
#require(summrt)
```

# Load the data

The evaluation data set contains synthetic data describing a simulated oubtreak for epi model cross validation. The ***all_data*** components used in this example include:

-   ***incubation:*** daily weights declaring the distrubution of incubation periods
-   ***serial\$px:*** daily weights declaring the distribution of the serial interval
-   ***reporting delay:*** daily weights declaring the distribution of the reporting delay
-   **r*t\$Rt_calc*:** **ğ‘…ğ‘¡** curve used in the construction of the synthetic data
-   ***cases\$daily reports*****:** daily reported cases with synthetic reporting delays

**all_data** also contains additional attributes not directly usable by this model.

```{r}
url <- "https://raw.githubusercontent.com/cmilando/RtEval/main/all_data.RDS"
all_data <- readRDS(url(url))
```

# Prepare model data structures

In this section we convert the previously loaded input data into the formats expected by **EpiLPS**.

## Extract the discrete distribution of the serial interval

```{r}
si_spec <- Idist(probs = all_data$serial$Px)
```

## Prepare incidence data

Here we zero fill NaN rows.

```{r}
incidence = all_data$cases$daily_reports
which(is.na(incidence))
incidence[1] <- 0
```

# Run the model

## Fit Rt from incidence and serial interval

In this section we use Laplacian-P-Splines to compute **ğ‘…ğ‘¡** from the serial interval probability vector and daily incidence curve via the ***estimR*** method.

```{r}
LPSfit <- estimR(incidence = incidence, si = si_spec$pvec)
```

## Apply time shifts

Here we take the probability distributions for incubation time and reporting delays and aggregate and round them into the integer values expected by the model.

```{r}
INCUBATION_SHIFT = round(weighted.mean(x = all_data$incubation$Day, w = all_data$incubation$Px))

REPORTINGDELAY_SHIFT = round(weighted.mean(x = all_data$reporting_delay$Day, w = all_data$reporting_delay$Px))
```

# Plot model results

## Prepare plot data structure

In this section we are extracting the percentile ranges from the modeled ***LPSfit*** and joining them into the output format expected by **ggplot**.

```{r}
plot_data <- data.frame(
  package = "EpiLPS",
  date = all_data$cases$day - INCUBATION_SHIFT - REPORTINGDELAY_SHIFT,
  Rt_median = LPSfit$RLPS$Rq0.50,
  Rt_lb = LPSfit$RLPS$Rq0.025,
  Rt_ub = LPSfit$RLPS$Rq0.975
)
```

## Create plot

Here we plot the results of the modeled **ğ‘…ğ‘¡** curve and it's distribution against the synthetic curve.

```{r}
as_tibble(plot_data) %>%
  ggplot() +
  geom_hline(yintercept = 1, linetype = "11") +
  # *******
  # this is the true r(t), back-calculated
  geom_line(aes(x = Day, y = Rt_calc), data = all_data$rt) +
  # *******
  geom_ribbon(aes(x = date, ymin = Rt_lb, ymax = Rt_ub, fill = package),
              alpha = 0.25) +
  geom_line(aes(x = date, y = Rt_median, color = package)) +
  coord_cartesian(ylim = c(0, 5)) +
  xlab("Days") +
  ylab("Rt") +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 14)
  )
```

# Extract the outcomes using `summrt`

```{r}
#outcomes <- summrt(LPSfit)
```
