---
title: "Simple R(t) estimation evaluation comparison"
description: "An R(t) evaluation across multiple packages, fitting to simulated data, standardizing outputs, and evaluating outputs"
author: 
date: "2024-09-25"
output:
  bookdown::html_vignette2:
    fig_caption: yes
    code_folding: show
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{R(t) evaluation across packages: fit to simulated data with known reports}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

The goal of this tutorial is to use multiple R(t) packages to fit to a simple, simulated outbreak
with a known R(t), serial interval, generation interval, and reporting delay probability mass function
(PMF). We will use the `summrt` package to generate standardized outputs, plot results, and 
quantitatively evaluate the accuracy in R(t) estimation.

Eventually we will expand this to additional vignettes that will fit to and evaluate more complex
simulated and real datasets (this might require not evaluating just accuracy and reliability
in R(t) but also nowcasting and forecasting expected observations and comparing to 
the true observations). 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install.packages('EpiEstim', repos = c('https://mrc-ide.r-universe.dev', 'https://cloud.r-project.org'))
# devtools::install_github("EpiForeSITE/summrt")

library(EpiNow2)
library(EpiEstim)
library(rtestim)
library(EpiLPS)
require(summrt)
library(ggplot2)
library (tidyverse)
library(kableExtra)
source(here::here("R/get_mae.R"))
```

## Load the simulated data

Load the dataset we will be fitting the R(t) estimation packages
to in this vignette.

This will eventually use data from the `rtdata` package that will be documented and
will describe the specific epidemiological use case this
data scenario is meant to represent. In this case, we are going to fit to data 
on the number of reported cases, with a known discrete generation interval probability
mass function (PMF) and reporting delay PMF that are also provided as data.

One of the goals of this evaluation using the `summrt` package is to standardize date/time indexing. All R(t) estimates from each package should be lined
up such that the time vector returned alongside the estimates follows the same indexing.

```{r load-rtdata}
# We will eventually replace this with package data specific to the dataset. 
# E.g. this might be our baseline infections, onset, report data
url <- "https://raw.githubusercontent.com/cmilando/RtEval/main/all_data.RDS"
all_data <- readRDS(url(url))

ggplot(all_data$rt) + 
  geom_line(aes(x = Day, y = Rt)) +
  geom_hline(aes(yintercept = 1), linetype = "dashed") + 
  xlab("Day") + ylab("R(t)") + 
  scale_y_continuous(trans = "log") + 
  theme_bw() + ggtitle("Simulated R(t)")

ggplot(all_data$generation) +
  geom_bar(aes(x = Day, y = Px), stat = "identity") + 
  xlab("Day") + ylab("Generation interval PMF") +
  theme_bw() 

ggplot(all_data$reporting_delay) +
  geom_bar(aes(x = Day, y = Px), stat = "identity") +
  xlab("Day") + ylab("Reporting delay PMF") +
  theme_bw()

ggplot(all_data$cases) +
  geom_bar(aes(x = day, y = daily_reports), stat = "identity") +
  xlab("Day") + ylab("Reported cases") + 
  theme_bw()
```


## Use each package to estimate R(t)

Fit each of the packages to the dataset. See the package specific vignettes for more
of a walk through for the decisions made for each package. 


### EpiNow2
```{r EpiNow2, warning=FALSE, message=FALSE}

incidence_df = data.frame(
  date = lubridate::make_date(2020, 3, 19) + 1:nrow(all_data$cases),
  confirm = as.vector(all_data$cases$daily_reports)
)

gi_pmf <- NonParametric(pmf = all_data$serial$Px)
sym_report_delay_pmf <- NonParametric(pmf = all_data$reporting_delay$Px)
incubation_pmf <- NonParametric(pmf = all_data$incubation$Px)

start.time <- Sys.time()
EpiNow2_obj <- epinow(
  data = incidence_df,
  generation_time = generation_time_opts(gi_pmf),
  delays = delay_opts(incubation_pmf + sym_report_delay_pmf),
  backcalc = backcalc_opts(prior = 'reports'),
  rt = rt_opts(rw = 1),
  stan = stan_opts(chains = 4, cores = 4)
)
end.time <- Sys.time()

run.time.epinow <- end.time-start.time
```
### EpiEstim
```{r EpiEstim}
# reporting data with dates indexed by integers starting at 0
incidence_df <- data.frame(
  dates = all_data$cases$day,
  I = as.vector(all_data$cases$daily_reports)
)
colnames(incidence_df) <- c('dates', 'I')

# Serial interval from data PMF-has leading 0 and is vector of probabilities
si_distr <- as.matrix(all_data$serial$Px)
if (all_data$serial$Day[1] == 1) si_distr <- c(0, si_distr)
si_distr

# Estimate R DAILY
start.time <- Sys.time()
EpiEstim_obj <- EpiEstim::estimate_R(
  incid = incidence_df,
  method = "non_parametric_si",
  config = make_config(list(
    si_distr = si_distr,
    t_start = 2:nrow(incidence_df),
    t_end = 2:nrow(incidence_df)
  )),
  backimputation_window = 10
)
end.time <- Sys.time()

run.time.epiestim <- end.time-start.time

```

### rtestim
```{r rtestim}
rr <- 2:nrow(all_data$cases)
start.time <- Sys.time()
rtestim_obj <- cv_estimate_rt(
  observed_counts = all_data$cases$daily_reports[rr],
  x = all_data$cases$day[rr],
  delay_distn = all_data$serial$Px
)
end.time <- Sys.time()
run.time.rtestim <- end.time-start.time
```

### EpiLPS
```{r}
si_spec <- Idist(probs = all_data$serial$Px)
incidence = all_data$cases$daily_reports
which(is.na(incidence))
incidence[1] <- 0
start.time <- Sys.time()
EpiLPS_obj <- estimR(incidence = incidence, si = si_spec$pvec)
end.time <- Sys.time()
run.time.epilps <- end.time-start.time
```

## Call the `SummRt` package to standardize the outputs 
```{r get-standardized-outputs}

std_EpiEstim <- summarize_rtestimate(EpiEstim_obj)
std_EpiNow2 <- summarize_rtestimate(EpiNow2_obj)
std_rtestim <- summarize_rtestimate(rtestim_obj)
std_EpiLPS <- summarize_rtestimate(EpiLPS_obj)

# Put them all together into one dataframe
convert_to_df <- function(output_summrt){
  df <- tibble::tibble(output_summrt$estimates) |>
  dplyr::mutate(package = output_summrt$package)
}

df <- tibble::tibble() |>
  bind_rows(convert_to_df(std_EpiNow2)) |>
  bind_rows(convert_to_df(std_rtestim)) |>
  bind_rows(convert_to_df(std_EpiLPS)) |>
  bind_rows(convert_to_df(std_EpiEstim)) |>
  dplyr::left_join(
    all_data$rt,
    by = c("date" = "Day")
  )

```

## Plot outputs 
We're going to make a plot of the four R(t) estimates that result from
the `summRt` output standardization, overlaid onto one another. 
Eventually we will write functionality to combine these things nicely and
plot them a few different ways. For now will just show how to do this
via accessing the elements directly.
```{r plot-outputs}

ggplot(data = df) +
    geom_line(aes(x = date,
                y = Rt),
            linetype = "dashed") +
  geom_line(
    aes(x = date,
                y = median,
                color = package))+
  geom_ribbon(
    aes(x = date,
                  ymin = lb,
                  ymax = ub,
              fill = package),
              alpha = 0.1) +
    scale_y_continuous(trans = "log10",
                     limits = c(0.1, 10)) +
    theme_bw() +
    xlab('Time (days)') +
    ylab('R(t)')


# Also plot them faceted
ggplot(data = df) +
    geom_line(aes(x = date,
                y = Rt),
            linetype = "dashed") +
  geom_line(
    aes(x = date,
                y = median,
                color = package))+
  geom_ribbon(
    aes(x = date,
                  ymin = lb,
                  ymax = ub,
              fill = package),
              alpha = 0.1) +
    scale_y_continuous(trans = "log10",
                     limits = c(0.1, 10)) +
    facet_wrap(~package) + 
    theme_bw() +
    xlab('Time (days)') +
    ylab('R(t)')

```
  

## Score the output of the R(t) model
Run times and MAE for each model
```{r score-output}
mae.epinow <- get_mae(pd = std_EpiNow2,                                                   rt_sim=all_data$rt)
mae.epiestim <- get_mae(pd = std_EpiEstim,                                                   rt_sim=all_data$rt)
mae.rtestim <- get_mae(pd = std_rtestim,                                                   rt_sim=all_data$rt)
mae.epilps <- get_mae(pd = std_EpiLPS,                                                   rt_sim=all_data$rt)

run.time.obj <- data.frame(Package=c("EpiNow2","EpiEstim","rtestim","EpiLPS"),
                Run.Time=round(c(run.time.epinow,
                                 run.time.epiestim,
                                 run.time.rtestim,
                                 run.time.epilps),3),
                MAE=round(c(mae.epinow,
                            mae.epiestim,
                            mae.rtestim,
                            mae.epilps),3))

run.time.obj %>%
  kbl() %>%
  kable_styling()

```


## Discussion/Interpretation
Which packages perform best on this dataset in what regimes? 
