---
title: "SummRt"
description: "An R(t) evaluation across multiple packages, fitting to simulated data, standardizing outputs, and evaluating outputs"
author: "Kaitlyn Johnson"
date: "2024-09-25"
output:
  bookdown::html_vignette2:
    fig_caption: yes
    code_folding: show
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{R(t) evaluation across packages: fit to simulated data with known reports}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

The goal of this tutorial is to use multiple R(t) packages to fit to a simple, simulated outbreak
with a known R(t), serial interval, generation interval, and reporting delay probability mass function
(PMF). We will use the `summrt` package to generate standardized outputs, plot results, and 
quantitatively evaluate the accuracy in R(t) estimation.

Eventually we will expand this to additional vignettes that will fit to and evaluate more complex
simulated and real datasets (this might require not evaluating just accuracy and reliability
in R(t) but also nowcasting and forecasting expected observations and comparing to 
the true observations). 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(EpiNow2)
library(EpiEstim)
library(rtestim)
library(EpiLPS)
#library(SummRt)
library(ggplot2)
library (tidyverse)
```

## Load the simulated data

Load the dataset we will be fitting the R(t) estimation packages
to in this vignette.

This will eventually use data from the `rtdata` package that will be documented and
will describe the specific epidemiological use case this
data scenario is meant to represent. In this case, we are going to 
fit to data on the number of reported
```{r load-rtdata}
# We will eventually replace this with package data specific to the dataset. 
# E.g. this might be our baseline infections, onset, report data
url <- "https://raw.githubusercontent.com/cmilando/RtEval/main/all_data.RDS"
all_data <- readRDS(url(url))
```


## Use each package to estimate R(t)

Fit each of the packages to the dataset. 
```{r fit-Rt-with-each-package}

#EpiNow2
source("02_EPINOW2_infections.R")

```
This will call `SummRt` package
functions which will use S3 methods to wrap around the individual 
R packages. All outputs will be in a standardized format that we can use to
evaluate the R(t) estimation (either from directly comparing to the 
true simulated R(t) or via comparing to an evaluation dataset of held out
retrospective observations).
```{r plot-output}
#source file, later will call function
source("03_plot.R")
plot_epinow2 <- this_plot(pd = plot_data, 
                          title = "title")
```
## Score the output of the R(t) model
This will either be compared to ground truth simulated data or
observations 
```{r score-output}
```

## Compare scores from each package
Visual comparison of scores for each package
```{r compare-scores-across-packages}
```

## Discussion/Interpretation
Which packages perform best on this dataset in what regimes? 
E.g. EpiEstim most accurately captures perturbation,
EpiNow2 performs best at nowcasting/forecasting observations
