---
title: "Rt Estimation with rtestim Package"
author: "Peiwen Liu, Abby Leung"
date: "2024-09-25"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rt Estimation with rtestim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r init, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Introduction

This vignette demonstrates how to use the `rtestim` package to estimate the reproduction number \(R_t\) for a disease based on observed case counts and other relevant delay distributions. We will load in pre-processed data, run the `estimate_rt` model, and visualize the results along with confidence bands.

## Load Data

To start, we will load the required data. The data includes daily reported cases, serial interval, incubation period, and reporting delay distributions.

```{r setup, echo=TRUE, include=TRUE, eval=TRUE}

library(rtestim)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(summrt)
url <- "https://raw.githubusercontent.com/cmilando/RtEval/main/all_data.RDS"
all_data <- readRDS(url(url))

```
The data includes:

+ daily_reports: Daily reported case counts
+ serial: Distribution of the serial interval between cases
+ incubation: Incubation period distribution
+ reporting_delay: Delay in reporting case data

These will be used to estimate 
  over the observation period.

## Running the Rt Estimation Model
Next, we will estimate the reproduction number \(R_t\) using the `estimate_rt` function. 

This function requires the observed daily case counts, the corresponding days, and the serial interval distribution as inputs.
rr defines the range of rows from the dataset all_data$cases that will be included in the     analysis.
```{r runcode, echo=TRUE, include=TRUE, eval=TRUE}
rr <- 2:nrow(all_data$cases)

rtestim <- cv_estimate_rt(
  observed_counts = all_data$cases$daily_reports[rr],
  x = all_data$cases$day[rr],
  delay_distn = all_data$serial$Px
)
```

## Extract the outomes using `summrt`

The estimate_rt function uses the daily reports and the serial interval to produce time-varying estimates of the reproduction number \(R_t\)Calculating Confidence Bands

We can compute approximate confidence bands around the \(R_t\)estimates using the confband function. These confidence bands help provide uncertainty intervals for the estimates.

``` {r approximate confidence bands, echo=TRUE, include=TRUE, eval=TRUE}
rtestim_cb <- confband(rtestim, lambda = "lambda.1se")
```

The lambda value, derived from cross-validation, controls the smoothness of the estimated curve. The function outputs the fit and credible intervals (e.g., 2.5% and 97.5% quantiles), which are used to visualize the uncertainty.

# Plotting output
We first compute the shifts for incubation and reporting delay.
``` {r shift, echo=TRUE, include=TRUE, eval=TRUE}
INCUBATION_SHIFT = round(weighted.mean(x = all_data$incubation$Day,
                                       w = all_data$incubation$Px))

REPORTINGDELAY_SHIFT = round(weighted.mean(x = all_data$reporting_delay$Day,
                                           w = all_data$reporting_delay$Px))
```

### create rtestim output
Adjust infection date based on incubation period and reporting delay period.
```{r df, echo=TRUE, include=TRUE, eval=TRUE}
plot_rtestim <- data.frame(
  package = "rtEstim",
  date = all_data$cases$day[rr] - INCUBATION_SHIFT - REPORTINGDELAY_SHIFT,
  Rt_median = rtestim_cb$fit,
  Rt_lb = rtestim_cb$`2.5%`,
  Rt_ub = rtestim_cb$`97.5%`
)
```

### plot format
``` {r plot, echo=TRUE, include=TRUE, eval=TRUE, fig.height = 4, fig.width = 7}
as_tibble(plot_rtestim) %>%
  ggplot() +
  geom_hline(yintercept = 1, linetype = "11") +
  # *******
  # this is the true r(t), back-calculated
  geom_line(aes(x = Day, y = Rt_calc), data = all_data$rt) +
  # *******
  geom_ribbon(aes(x = date, ymin = Rt_lb, ymax = Rt_ub, fill = package), alpha = 0.25) +
  geom_line(aes(x = date, y = Rt_median, color = package)) +
  coord_cartesian(ylim = c(0, 5)) +
  xlab("Days") +
  ylab("Rt") +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 14)
  )
```
