---
title: "Rt Estimation with rtestim Package"
author: "Peiwen Liu, ..."
date: "2024-09-25"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rt Estimation with rtestim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette demonstrates how to use the `rtestim` package to estimate the reproduction number \(R_t\) for a disease based on observed case counts and other relevant delay distributions. We will load in pre-processed data, run the `estimate_rt` model, and visualize the results along with confidence bands.

# Load Data

To start, we will load the required data. The data includes daily reported cases, serial interval, incubation period, and reporting delay distributions.

```{r setup, echo=TRUE, include=TRUE, eval=TRUE}
library(rtestim)
url <- "https://raw.githubusercontent.com/cmilando/RtEval/main/all_data.RDS"
all_data <- readRDS(url(url))

```


# Run the model
```{r runcode, echo=TRUE, include=TRUE, eval=TRUE}
rr <- 2:nrow(all_data$cases)

rtestim <- estimate_rt(
  observed_counts = all_data$cases$daily_reports[rr],
  x = all_data$cases$day[rr],
  delay_distn = all_data$serial$Px
)

# Approximate confidence bands
rtestim_cb <- confband(rtestim, lambda = rtestim$lambda[37])
```

# Plotting output
We first compute the shifts for incubation and reporting delay.
``` {r shift, echo=TRUE, include=TRUE, eval=TRUE}
INCUBATION_SHIFT = round(weighted.mean(x = all_data$incubation$Day,
                                       w = all_data$incubation$Px))

REPORTINGDELAY_SHIFT = round(weighted.mean(x = all_data$reporting_delay$Day,
                                           w = all_data$reporting_delay$Px))
```
Create dataframe needed 
``` {r plot, echo=TRUE, include=TRUE, eval=TRUE}
as_tibble(plot_rtestim) %>%
  ggplot() +
  geom_hline(yintercept = 1, linetype = "11") +
  # *******
  # this is the true r(t), back-calculated
  geom_line(aes(x = Day, y = Rt_calc), data = all_data$rt) +
  # *******
  geom_ribbon(aes(x = date, ymin = Rt_lb, ymax = Rt_ub, fill = package), alpha = 0.25) +
  geom_line(aes(x = date, y = Rt_median, color = package)) +
  coord_cartesian(ylim = c(0, 5)) +
  xlab("Days") +
  ylab("Rt") +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 14)
  )

saveRDS(plot_rtestim, file = 'plot_objects/plot_data_rtestim_reports.RDS')
```
# Extract the outomes using `summrt`
